Type Ia Supernova Explosions from Hybrid Carbon-Oxygen-Neon White Dwarf Progenitors

### 2017-10-14 paper/ms.tex
Revert "Added citation to Willcox et al. 2016."
This reverts commit 7a7af483f4f88efe46a18823ca013f26741dc05d.
Revert "First pass at the intro."
This reverts commit 203b654688ab677ddc6c83c153f75c4e2157f548.

\author{
Carlyn N.\ Augustine\altaffilmark{1},
Donald E.\ Willcox\altaffilmark{2},
Dean M.\ Townsley\altaffilmark{3},
and Alan C.\ Calder\altaffilmark{2,4}
}

\altaffiltext{1}{
  Department of Physics and Astronomy,
  The University of Alabama, Tuscaloosa, AL, 35487-0324, USA
}
\altaffiltext{2}{
  Department of Physics and Astronomy,
  Stony Brook University, Stony Brook, NY, 11794-3800, USA; \\
  \href{mailto:donald.willcox@stonybrook.edu}{donald.willcox@stonybrook.edu}
}
\altaffiltext{3}{
  Institute for Advanced Computational Sciences,
  Stony Brook University, Stony Brook, NY, 11794-5250, USA
}

\begin{abstract}
We present simulations of thermonuclear (type Ia) supernovae
from hybrid C/O/Ne progenitor white dwarfs. These progenitors 
were evolved though accretion with the \MESA\ stellar evolution code 
and include blah blah about the new hybridocity....
For comparison, we performed a set of supernova simulations from
traditional C/O progenitors....
We find ....
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:intro}
Type Ia supernovae (\SNeIa) are bright stellar explosions....

Phillips relation
~\citep{phillips:absolute}.

Cosmology
due to dark energy \citep{riess.filippenko.ea:observational,
perlmutter.aldering.ea:measurements,leibundgut2001}.

Brief discussion of progentors...
ancient history~\citep{hoylefowler60,arnett.truran.ea:nucleosynthesis}. The
Single degenerate 
\citep{hoylefowler60,trucam71,whelaniben73,Nomo84}. Alternately, in

Other models
\citep{webbink84,ibentutukov84}, in which two WDs inspiral and merge.

Hybrid models 
old 
\citep{denissenkovetal2015}
\citet{Wangetal2014} and \citet{Mengetal2014} presented binary

Recent results from stellar evolution
The models we are exploding \citep{brooksetal2017}

If we refer to the work by name, it should appear ``The hybrid models
of~\citet{brooksetal2017}." Or, we can just cite it~\citep{brooksetal2017}.

@@

Explain the Single Degenerate scenario
\citep{Baraffe2004Stability-of-Su,WoosWunsKuhl04,wunschwoosley2004,Kuhletal06,nonakaetal2012}.
expansion \citep{Nomo84,WoosWunsKuhl04}, and a flame is born.

Explain DDT

Go through below and cite some of these as you tell the story

d~\citep{arnett.truran.ea:nucleosynthesis}.
~\citep{mazzalietal2008}.

\citep{khokhlov1993,bychovliberman1995,
SNrt,Khok95,NiemHill95,khoketal1997,ZingDurs07,
cholazarianvishniac2003,
roepkehn2003,roepkehn2004,
Zingale2005Three-dimension,Schmetal06a, Schmetal06b,Aspdetal08,
Woosetal09,csetal2009,hicksrosner2013,c-ssr2013,
jacketal2014,poludnenko2015,hicks2015}.

A deflagration alone will not produce a event of normal brightness and
expansion velocity~\citep{roepkeetal07}. Instead, the initial
deflagration must transition to a detonation after the star has
expanded some in order to produce abundances and a stratified ejecta
in keeping with
observations~\citep{Khokhlov1991Delayed-detonat,hokowh95}.
The physics of this ``deflagration-to-detonation transition'' (DDT)
are not completely understood, but there has been considerable study
based on mechanisms involving flame fronts in highly turbulent
conditions~\citep{1986SvAL, woosley90, Khokhlov1991Delayed-detonat,
  hokowh95, HoefKhok96, khoketal1997, NiemWoos97, hwt98, Niem99,
  GameKhokOran05,roepke07, poletal2011,c-ssr2013,poludnenko2015}.
These models generally reproduce the observations under certain
assumptions about the ignition~\citep{townetal2009}, but research has
shown that the results are very sensitive to the details of the
ignition~\citep{PlewCaldLamb04,GameKhokOran05,garciasenz:2005,
  roepkeetal07,Jordan2008Three-Dimension}.

Summarize what we do a la:
In our simulations, we
initialize a detonation once the deflagration front reaches a
characteristic DDT fuel density, which controls the degree of
expansion the star undergoes during the deflagration stage. The
implementation details are described further in
\secref{subsec:ddt_details}.

>>>>

\author{
Donald E.\ Willcox\altaffilmark{1},
Dean M.\ Townsley\altaffilmark{2},
Alan C.\ Calder\altaffilmark{3,4},
Pavel A.\ Denissenkov\altaffilmark{5,6},
and Falk Herwig\altaffilmark{5,6}
}

\altaffiltext{1}{
  Department of Physics and Astronomy,
  Stony Brook University, Stony Brook, NY, 11794-3800, USA; \\
  \href{mailto:donald.willcox@stonybrook.edu}{donald.willcox@stonybrook.edu}
}
\altaffiltext{2}{
  Department of Physics and Astronomy,
  The University of Alabama, Tuscaloosa, AL, 35487-0324, USA
}
\altaffiltext{3}{
  Department of Physics and Astronomy,
  Stony Brook University, Stony Brook, NY, 11794-3800, USA
}
\altaffiltext{4}{
  Institute for Advanced Computational Sciences,
  Stony Brook University, Stony Brook, NY, 11794-5250, USA
}
\altaffiltext{5}{
  Department of Physics and Astronomy, University of Victoria, PO Box 1700, STN CSC, 
  Victoria, BC V8W 2Y2, Canada
}
\altaffiltext{6}{
  Joint Institute for Nuclear Astrophysics, Notre Dame, IN 46556, USA
}

\begin{abstract}
Motivated by recent results in stellar evolution that predict the
existence of hybrid white dwarf (WD) stars with a C-O core inside an
O-Ne shell, we simulate thermonuclear (Type Ia) supernovae from these
hybrid progenitors. We use the \FLASH\ code to perform
multidimensional simulations in the deflagration to detonation
transition (DDT) explosion paradigm. Our hybrid progenitor models were
produced with the \MESA\ stellar evolution code and include the
effects of the Urca process, and we map the progenitor model to the
\FLASH\ grid. We performed a suite of DDT simulations over a range of
ignition conditions consistent with the progenitor's thermal and
convective structure assuming multiple ignition points. To compare the
results from these hybrid WD stars to previous results from C-O white
dwarfs, we construct a set of C-O WD models with similar properties
and similarly simulate a suite of explosions.  We find that despite
significant variability within each suite, trends distinguishing the
explosions are apparent in their \Ni{56} yields and the kinetic
properties of the ejecta. We compare our results to other
  recent work that studies explosions from these hybrid progenitors.
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:intro}
Type Ia supernovae (\SNeIa) are bright stellar explosions that 
produce $\sim 0.6~\Msun$ of radioactive \Ni{56}, the decay of
which powers the light curve and leads to a relation between
the peak brightness of an event and the rate of its decline
from maximum~\citep{phillips:absolute}. This relation
enables \SNeIa\ to be used as ``standard
candles'' for cosmological studies, and this use led to 
the discovery that the expansion of the Universe is accelerating 
due to dark energy \citep{riess.filippenko.ea:observational,
perlmutter.aldering.ea:measurements,leibundgut2001}.

Despite intense study (driven in part by their application as distance
indicators for cosmology), we still have only an incomplete
understanding of the explosion mechanism and fundamental questions,
such as the likely progenitor system(s), persist. It is widely
accepted that \SNeIa\ result from the thermonuclear explosion of a
white dwarf (WD) composed largely of C and O, with this understanding
going back many
decades~\citep{hoylefowler60,arnett.truran.ea:nucleosynthesis}. The
rapid thermonuclear fusion of C and O in a WD releases enough energy
to unbind it, produces the \Ni{56} necessary to power the light curve,
and explains the lack of H observed in the spectra.

There are, however, several possible progenitor systems for such a
configuration. All models involve a binary system and at least one C-O
WD, which follows from the original association of \SNeIa\ with C-O
burning under degenerate conditions~\citep{hoylefowler60}.
%awk
Some proposed systems posit a single white dwarf, the single
degenerate (SD) paradigm, and some posit the merger or collision of
two white dwarfs, the double degenerate (DD) paradigm, and within
these are variations.
%\citep{hoylefowler60,webbink84,ibentutukov84}.

The ``classic'' model is the Chandrasekhar-mass model in which a white
dwarf gains mass from a companion, a main sequence or red giant star,
or perhaps a He WD, and a thermonuclear runaway occurs just as it
approaches the Chandrasekhar limiting mass
\citep{hoylefowler60,trucam71,whelaniben73,Nomo84}. Alternately, in
the sub-Chandrasekhar-mass scenario, explosive burning in the accreted
layer triggers a detonation at the surface or in the core of a
lower-mass WD \citep{nomoto80,woosleyweavertaam80,simetal2010}.

The other main class of models is the DD progenitor,
\citep{webbink84,ibentutukov84}, in which two WDs inspiral and merge.
This scenario includes inspiraling pairs, collisions, violent mergers,
and also the ``core-degenerate'' model where the merger takes place in
a common envelope
\citep{raskinetal2009,pakmoretal2011,kashi:2011,pakmoretal2012a,Shenetal12}
Also
see~\citet{hillebrandt.niemeyer:type,howell2011,hillebrandtetal2013,calderetal2013}
for additional discussion.

The observational evidence of one progenitor vs.\ another is
conflicting. \SNeIa\ show a wide range of luminosities and also the
possibility that there are two classes of
progenitor~\citep{MannucciEtAl06,howelletal+09,howell2011}. Observational
and population synthesis~\citep{yungelsonlivio2000, Maoz2008,
  Toonenetal2012, Claeysetal2014} arguments suggest that there simply
may not be enough Chandrasekhar-mass progenitors to explain the
observed \SNIa\ rate.  There is, however, disagreement over the
significance of these observations~\citep{hachisu:2008} and our
understanding of the different progenitor systems is subject to
uncertainties in aspects of their pre-supernova evolution, including
the effects of self-absorption on the number of supersoft X-ray
sources predicted from the SD
channel~\citep{Nielsenetal2013}. Certainly there is disagreement in
the interpretation of observations that stems from uncertainty in the
accretion phase of SD evolution~\citep{hachisu:2010}.

Additionally, the oft-cited claim that the WD in the SD channel would
lose mass via nova explosions, thereby preventing it from reaching the
Chandrasekhar mass, is questioned on the basis of high observed WD
masses in cataclysmic variables (CVs) relative to
pre-CVs~\citep{Zorotovic:2011}. Prior work on novae and rapidly
accreting WDs strongly suggests that novae will not be able to
grow~\citep{denissenkovetal2013b,denissenkovetal2014}, especially not
from WD masses of $0.83~\Msun$, suggested to be the mean mass of WDs
in CVs in~\citet{Zorotovic:2011}. Recently, binary population
synthesis calculations by~\citet{Wijnenetal2015} explored both thermal
time-scale mass transfer and mass increase during novae as potential
solutions, finding that both mechanisms result in contradictions with
observations and cannot satisfy the observed WD mass difference.

However, if it is possible to get a \SNeIa\ out of the SD scenario,
then ``hybrid'' C-O-Ne WDs (WDs with a C-O core in an O-Ne
shell)~\citep{denissenkovetal2015} may play a key role (cf.
\secref{section:hybrid_progenitor_models}). Prior stellar evolution
work by~\citet{GarciaBerroetal1997} produced an O-Ne WD with a \C{12}
abundance as high as $0.048$ and suggested that this is sufficient to
ignite an explosion if the WD accreted matter to approach the
Chandrasekhar mass. The hybrid WD of~\citet{denissenkovetal2015} has a
significantly higher average \C{12} abundance of $0.17$ and such WDs
would provide readily ignitable Ia progenitors that are already very
close to the Chandrasekhar limiting mass, and are therefore perhaps
the most likely to produce a \SNeIa.

Acknowledging the ignitability of hybrid C-O-Ne WDs,
\citet{Wangetal2014} and \citet{Mengetal2014} presented binary
population synthesis studies for these hybrid WDs and suggested they
find consistency with the observed properties of Type Iax events,
though uncertainties remain in estimating the rate of \SNeIa\ from
hybrid WDs due to the uncertain carbon-burning rate and
common-envelope ejection efficiency. The study we present here
addresses \SNeIa\ explosions from these hybrid models and other
research groups have also recently explored this possiblilty using
hydrodynamic simulations.  \citet{kromeretal2015} performed pure
deflagration simulations from C-core models and \citet{bravoetal2016}
simulated explosions from a variety of progenitor models, including
WDs with similar stellar evolution to those of
\citet{denissenkovetal2015}.  We compare these explosion studies to
ours in the discussion below.

While there is uncertainty, some contemporary observations do strongly
support the SD progenitor. Events like PTF11kx and others show distinct 
circumstellar shells of material that can be best explained in the SD
context~\citep{dildayetal2012,silverman:2013}. The SNR 3C 397 is heralded
as a case where only an explosion from a Chandrasekhar-mass progenitor can 
produce the nuclei seen in the remnant, due to the need for electron captures 
at high density~\citep{3c397}.  
The recent observation of a UV pulse~\citep{cao:2015,cao:2015e} in the
early evolution of an \SNIa\ also supports the SD model. 
Observations of remnants also offer support 
for Chandrasekhar-mass explosions, including wind-blown shells in 
RCW86~\citep{williams:2011} and shocked circumstellar material/bubble in the Kepler
remnant~\citep{chiotellis:2012,burkey:2013}.  Altogether, there is
substantial evidence that suggests that the SD channel plays a role in
at least some of the observed \SNeIa~\citep{baron2014}.

The sub-Chandrasekhar-mass model does not have the population
synthesis arguments working against it and we know low-mass WDs in
binary systems exist. Systems that are believed will evolve to an
explosion have been observed~\citep{kilic:2014}, potential events have
been identified~\citep{geier:2013,inserra:2015}, and the Type Iax
sub-class of SN Ia~\citep{foleyetal2013,wang:2013} have been suggested
as being sub-Chandrasekhar-mass events themselves.

Observational evidence also supports the DD progenitor system, and
the scenario is  increasingly seen as the likely progenitor of some
events. SN 2011fe has been intensely observed and does not show features
in its spectra that would be expected if there were a normal stellar
companion~\citep{graham:sn2011fe}, suggesting a DD system. Super-Chandrasekhar
mass explosions like SN~2007if~\citep{scalzo:2010,Yuan:2010} and
SNLS~03D3bb~\citep{howell+06} also suggest mergers. There are
also many population synthesis arguments in favor of mergers as well 
\citep[see][for a review]{maoz:2014}.

@@

In the Chandrasekhar-mass scenario, the central temperature and 
density of the WD increase as it accretes mass from a binary companion 
and approaches the limiting Chandrasekhar mass. As the mass
approaches the limit, central conditions become hot enough for 
carbon fusion to begin (via the \C{12}--\C{12} reaction), driving 
the development of convection throughout the interior of the WD
\citep{Baraffe2004Stability-of-Su,WoosWunsKuhl04,wunschwoosley2004,Kuhletal06,nonakaetal2012}.
As the central temperature reaches $\sim 7\times 10^8~\mathrm{K}$, the
fuel in a convective plume burns to completion before it can cool via
expansion \citep{Nomo84,WoosWunsKuhl04}, and a flame is born.

The nature of this burning, be it a supersonic detonation or subsonic 
deflagration, largely determines the outcome of the explosion. 
It has been known for some time that a purely supersonic burning front 
cannot
explain observations because the supersonic front very rapidly incinerates 
the star without it having time to react and 
expand~\citep{arnett.truran.ea:nucleosynthesis}. The lack of expansion 
allows most of the star to burn at high densities,
which produces an excessive $^{56}$Ni yield and does not match the
stratified composition of observed remnants~\citep{mazzalietal2008}.  
Instead, a subsonic deflagration must ignite, which allows the
outer layers of the star to expand ahead of the burning front.
In this case, the density of the expanding material decreases, 
which leads to incomplete burning of more mass and thus
increased production of intermediate mass elements. This
deflagration must accelerate via instabilities and turbulent
interaction, a topic that has been explored extensively in the 
past~\citep{khokhlov1993,bychovliberman1995,
SNrt,Khok95,NiemHill95,khoketal1997,ZingDurs07,
cholazarianvishniac2003,
roepkehn2003,roepkehn2004,
Zingale2005Three-dimension,Schmetal06a, Schmetal06b,Aspdetal08,
Woosetal09,csetal2009,hicksrosner2013,c-ssr2013,
jacketal2014,poludnenko2015,hicks2015}.

A deflagration alone will not produce a event of normal brightness and
expansion velocity~\citep{roepkeetal07}. Instead, the initial
deflagration must transition to a detonation after the star has
expanded some in order to produce abundances and a stratified ejecta
in keeping with
observations~\citep{Khokhlov1991Delayed-detonat,hokowh95}.
The physics of this ``deflagration-to-detonation transition'' (DDT)
are not completely understood, but there has been considerable study
based on mechanisms involving flame fronts in highly turbulent
conditions~\citep{1986SvAL, woosley90, Khokhlov1991Delayed-detonat,
  hokowh95, HoefKhok96, khoketal1997, NiemWoos97, hwt98, Niem99,
  GameKhokOran05,roepke07, poletal2011,c-ssr2013,poludnenko2015}.
These models generally reproduce the observations under certain
assumptions about the ignition~\citep{townetal2009}, but research has
shown that the results are very sensitive to the details of the
ignition~\citep{PlewCaldLamb04,GameKhokOran05,garciasenz:2005,
  roepkeetal07,Jordan2008Three-Dimension}.
In our simulations, we
initialize a detonation once the deflagration front reaches a
characteristic DDT fuel density, which controls the degree of
expansion the star undergoes during the deflagration stage. The
implementation details are described further in
\secref{subsec:ddt_details}.
